---
title: "logitr"
output: github_document
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/logitr)](https://cran.r-project.org/package=logitr)

### ***This package is experimental and not regularly maintained. It is recommended to  use {glmnet} instead!

## Overview

`logitr` is an R package for linear and logistic regression with optional ridge and bridge regularization penalties. A (possibly incomplete) list of functions contained in the package can be found below:

* `logisticr()` computes the coefficient estimates for logistic regression (ridge and bridge regularization optional)
* `linearr()` computes the linear regression coefficient estimates (ridge regularization and weights optional)
* `predict.logisticr()` generates predictions and loss metrics for logistic regression
* `predict.linearr()` generates predictions and loss metric for linear regression

See [vignette](https://mgallow.github.io/logitr/) or [manual](https://github.com/MGallow/logitr/blob/master/logitr.pdf).

## Installation

```{r, eval = FALSE}
# The easiest way to install is from the development version from GitHub:
# install.packages("devtools")
devtools::install_github("MGallow/logitr")
```

If there are any issues/bugs, please let me know: [github](https://github.com/MGallow/logitr/issues). You can also contact me via my [website](http://users.stat.umn.edu/~gall0441/). Pull requests are welcome!

## Usage

```{r, message = FALSE}
library(logitr)

#we will use the iris data set
X = dplyr::select(iris, -c(Species, Sepal.Length))
y = dplyr::select(iris, Sepal.Length)
y_class = ifelse(dplyr::select(iris, Species) == "setosa", 1, 0)

#ridge regression (use CV for optimal lambda)
linearr(X, y, penalty = "ridge")

#ridge logistic regression (IRLS) (use CV for optimal lambda)
logisticr(X, y_class, penalty = "ridge")

#ridge logistic regression (MM)
logisticr(X, y_class, lam = 0.1, penalty = "ridge", method = "MM")

#bridge logistic regression (MM)
fit = logisticr(X, y_class, lam = 0.1, alpha = 1.2, penalty = "bridge")
fit

#predict using bridge logistic regression estimates
predict(fit, X[1:3,], y_class[1:3])

```
