---
title: "logitr"
output: github_document
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DOCUMENTATION IN PROGRESS***

See [vignette](https://htmlpreview.github.io/?https://github.com/MGallow/logitr/blob/master/Vignette.html) or [manual](https://htmlpreview.github.io/?https://github.com/MGallow/logitr/blob/master/logitr.pdf).

## Overview

`logitr` is an R package for logistic regression and ridge-penalized logistic regression. A (possibly incomplete) list of functions contained in the package can be found below:

* `logisticr()` computes the coefficient estimate for logistic regression. L2-penalization optional.
* `linearr()` computes the coefficient estimates for linear regression. L2-penalization and weighted linear regression, optional.
* `predict.logisticr()`
* `predict.linearr()`



## Installation

```{r, eval = FALSE}
# The easiest way to install is from the development version from GitHub:
# install.packages("devtools")
devtools::install_github("MGallow/logitr")
```

If there are any issues/bugs, please let me know: [github](https://github.com/MGallow/logitr/issues). You can also contact me via my [website](http://users.stat.umn.edu/~gall0441/).

## Usage

```{r, message = FALSE}
library(logitr)

#IN PROGRESS
```
