---
title: "logitr"
author: "Matt Galloway"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## DOCUMENTATION IN PROGRESS (longer version of README)***

## Overview

`logitr` is an R package for linear and logistic regression with optional ridge and bridge regularization penalties. A (possibly incomplete) list of functions contained in the package can be found below:

* `logistic()` calculate coefficient estimates for ridge (L2)-penalized linear regression
* `logistic_bridge()` is a bi-section search algorithm.



## Installation

```{r, eval = FALSE}
# The easiest way to install is from the development version from GitHub:
# install.packages("devtools")
devtools::install_github("MGallow/logitr")
```

If there are any issues/bugs, please let me know: [github](https://github.com/MGallow/logitr/issues). You can also contact me via my [website](http://users.stat.umn.edu/~gall0441/).

## Usage

```{r, message = FALSE}
library(logitr)

#IN PROGRESS
```
